---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cusumcharter

<!-- badges: start -->

[![R-CMD-check](https://github.com/johnmackintosh/cusumcharter/workflows/R-CMD-check/badge.svg)](https://github.com/johnmackintosh/cusumcharter/actions)

[![Codecov test coverage](https://codecov.io/gh/johnmackintosh/cusumcharter/branch/master/graph/badge.svg)](https://codecov.io/gh/johnmackintosh/cusumcharter?branch=master)

[![Render README](https://github.com/johnmackintosh/cusumcharter/actions/workflows/render-readme.yaml/badge.svg)](https://github.com/johnmackintosh/cusumcharter/actions/workflows/render-readme.yaml)

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/cusumcharter)](https://CRAN.R-project.org/package=cusumcharter)
<!-- badges: end -->

The goal of cusumcharter is to create both simple CuSum charts, with and without control limits from a vector, or to create multiple CuSum charts, with or without control limits, from a grouped dataframe, tibble or data.table

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("johnmackintosh/cusumcharter")
```

## A Simple CuSum calculation

This returns the CuSum statistics for a single vector, centred on a supplied target value:

```{r example}
library(cusumcharter)
test_vec <- c(0.175, 0.152, 0.15, 0.207, 0.136, 0.212, 0.166)

CuSum_res <- cusum_single(test_vec, target = 0.16)
CuSum_res
```

## Expanded outputs with cusum_single_df

This function takes a single vector as input and returns a data.frame with additional information used to calculate the CuSum statistic

```{r, example2}
test_vec2 <- c(0.175, 0.152, 0.15, 0.207, 0.136, 0.212, 0.166)
cusum_single_df(test_vec2, target = 0.16)
```

Here we don't supply a target, so one is calculated along with the other statistics
```{r}
test_vec3 <- c(1,1,2,11,3,5,7,2,4,3,5)
cusum_single_df(test_vec3)
```



## CuSum control limits

Two additional functions allow you to calculate control limits from a single vector and plot a CuSum chart with control limits. 

```{r, cusum_control_example}
test_vec3 <- c(1,1,2,3,5,7,11,7,5,7,8,9,5)
controls <- cusum_control(test_vec3, target = 4)
controls
```

## CuSum Control Chart

```{r single-control_chart, fig.width=5, fig.height=3,out.width= "100%"}
test_vec3 <- c(1,1,2,3,5,7,11,7,5,7,8,9,5)
controls <- cusum_control(test_vec3, target = 4)

cusum_control_plot(controls, 
                   xvar = obs, 
                   title_text = "CuSum out of control since 7th observation")
```


## Multiple CuSum Control Charts 

Using base R to split and apply our dataset, and ```data.table```'s ```rbindlist()``` to combine the results into a data.frame / data.table for  plotting. 

The x axis variable is the obervation number (```obs```) and the facet variable is ```.id```. Both these are generated by the ```cusum_control``` function.

```{r faceted_chart1, fig.width=5, fig.height= 3,out.width="100%"}
library(data.table)
library(ggplot2)
library(cusumcharter)

testdata <- data.frame(
  stringsAsFactors = FALSE,
  N = c(1L,2L,1L,3L,1L,1L,1L,1L,1L,
        1L,3L,2L,3L,2L,7L,11L,7L,9L),
  metric = c("metric1","metric1","metric1","metric1","metric1",
           "metric1","metric1","metric1","metric1","metric2",
           "metric2","metric2","metric2","metric2","metric2",
           "metric2","metric2","metric2"))

testlist <- split(testdata$N,testdata$metric)

testres <- lapply(testlist, cusumcharter::cusum_control)

testres <- data.table::rbindlist(testres,fill = TRUE, idcol = TRUE)

p <- cusum_control_plot(testres, xvar = obs, 
                        facet_var = .id, 
                        title_text = " faceted CuSum Control plots")
p

```

## Flexible x axis

Here we  add a date column, specify that the ```scale_type``` is ```'date'```, and provide the ```datebreaks``` argument to plot our data over time

```{r faceted_chart2, fig.width=5, fig.height= 3,out.width="100%"}
library(data.table)
library(ggplot2)
library(cusumcharter)

testdata <- data.frame(
  stringsAsFactors = FALSE,
 N = c(1L,2L,1L,3L,1L,1L,1L,1L,1L,
        1L,3L,2L,3L,2L,7L,11L,7L,9L),
  metric = c("metric1","metric1","metric1","metric1","metric1",
           "metric1","metric1","metric1","metric1","metric2",
           "metric2","metric2","metric2","metric2","metric2",
           "metric2","metric2","metric2"))

testlist <- split(testdata$N,testdata$metric)

testres <- lapply(testlist, cusumcharter::cusum_control)

testres <- data.table::rbindlist(testres,fill = TRUE, idcol = TRUE)
testres[, datecol := c("2021-01-01","2021-01-02", "2021-01-03", "2021-01-04" ,"2021-01-05", "2021-01-06","2021-01-07", "2021-01-08", "2021-01-09"), by = .id]

testres[,datecol := as.Date(datecol)]

p2 <- cusum_control_plot(testres, xvar = datecol, 
                         facet_var = .id, 
                         title_text = " faceted plots with date axis", 
                         scale_type = "date", 
                         datebreaks = '2 days')

p2 <- p2 + ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90,
                                                              hjust = 1, 
                                                              vjust = 0.5))
p2
```


